### 2011-11-07 doc/pldi2012/lowlevel.tex
lowlevel

The potential space concern with events is addressed by a {\em generation scheme}
that allows events to be reclaimed and reused without tracking references.  The
naive implementation of an event cannot reclaim the storage used by an event
until the event has triggered and all waiters on the event have been
notified.  
%Some of these waiters may take quite a while to query the event.
%Additionaly, we must somehow track how many queries to expect to know
%when they've all been performed.  
Our events include a
{\em generation count} that tracks how many times the event has been triggered,
which allows an event to ve reused with a new generation as soon as the previous generation has triggered.
Holders of an older (handle,generation) pair are able to query the newer
incarnation of the event and determine from the generation mismatch that their
event must have already triggered.  With this technique, there is no need to
count references to the event at all. 
% An event can only be reused until the
%generation counter reaches its maximum value, the current implementation
%supports $2^{32}-1$ generations, and should that not be sufficient, extending
%it to $2^{64}-1$ is trivial.

@@

The deferred nature of a deferred lock also allows a distributed implementation
of locks.  Similar to events, a deferred lock can be created on any node
without communication with the other nodes.  That node becomes
the initial home for the lock, but the deferred lock (unlike events) can 
migrate to other nodes or even be shared between nodes, allowing repeated
lock requests from a single node to be kept local to that node after a single
message is sent to the current owner.  (Should the ownership of the lock change
while the message is in flight, the previous owner forwards the request to
the new owner.  The highly unlikely, but possible, case in which multiple 
forwardings are required is bounded at the number of nodes in the machine.)
The ability to migrate locks between nodes allows the performance of a
heavily-contended cluster-level lock to approach that of a single-node lock.

>>>>

The potential space concern with events is addressed by a {\em generation scheme}
that allows events to be reclaimed and reused without tracking references.  The
naive implementation of an event cannot reclaim the storage used by an event
until the event has triggered and all waiters on the event have been
notified.  
%Some of these waiters may take quite a while to query the event.
%Additionaly, we must somehow track how many queries to expect to know
%when they've all been performed.  
Our events include a
{\em generation count} that tracks how many times the event has been triggered,
which allows an event to be reused with a new generation as soon as the previous generation has triggered.
Holders of an older (handle,generation) pair are able to query the newer
incarnation of the event and determine from the generation mismatch that their
event must have already triggered.  With this technique, there is no need to
count references to the event at all. 
% An event can only be reused until the
%generation counter reaches its maximum value, the current implementation
%supports $2^{32}-1$ generations, and should that not be sufficient, extending
%it to $2^{64}-1$ is trivial.

@@

The deferred nature of a deferred lock also allows a distributed implementation
of locks.  Similar to events, a deferred lock can be created on any node
without communication with the other nodes.  That node becomes
the initial home for the lock, but the deferred lock (unlike events) can 
migrate to other nodes or even be shared between nodes, allowing repeated
lock requests from a single node to be kept local to that node after a single
message is sent to the current owner.  
%(Should the ownership of the lock change
%while the message is in flight, the previous owner forwards the request to
%the new owner.  The highly unlikely, but possible, case in which multiple 
%forwardings are required is bounded at the number of nodes in the machine.)
The ability to migrate locks between nodes allows the performance of a
heavily-contended cluster-level lock to approach that of a single-node lock.
